### 2018-11-05 evoforces_cletalk/slides.Rmd
small edits to slides

---
title: "Challenges in<br>detecting evolutionary forces<br>in language change <br>using diachronic corpora"

author: "Andres Karjus<br><font size=5>(supervised by Kenny Smith, Richard A. Blythe, Simon Kirby)<br>Centre for Language Evolution, University of Edinburgh</font>"
date: "CLE seminar, 6.11.2018"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: false
      countIncrementalSlides: false
      ratio: "16:9"
      navigation:
        scroll: false
---
class:inverse

<style>
.remark-slide-content {
  padding-top: 7px;
  padding-left: 25px;
  padding-right: 20px;
  padding-bottom: 30px;
}
body { 
  line-height: 3em;
} 
.mjx-chtml{ font-size: 100% !important; } 
.small { font-size: 50%; margin-top:0em; margin-bottom:0em}

p {margin-bottom:0em}

</style>

```{r setup, echo=F}
options(htmltools.dir.version = FALSE)
options(servr.daemon = TRUE)
knitr::opts_chunk$set(echo = FALSE, message=F,warning=F,dev="png",dpi=100)
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_light(
  base_color = rgb(0,0,0.2),
  black_color = "black",
  background_color = "white",
  header_font_google = google_font("Nanum Gothic"),
  text_font_google   = google_font("Nanum Gothic", "300", "300i"),
  text_slide_number_font_size = "0.4em",
  header_h1_font_size = "50px",
  text_font_size = "35px",
  text_font_family_fallback = "Helvetica",
  code_font_google   = google_font("Droid Mono"),
  title_slide_background_image = "cle.png",
  title_slide_background_size = "8%",
  title_slide_background_position = "top right",
  link_color = "lightblue"
)
```

```{r functions}
#knitr::knit_exit() # titleslide debug handbrake

library(RColorBrewer)
library(dplyr)
library(plotly)
load("/Users/pplsuser/Dropbox/phd/selection_replypaper/cletalk/binresults.RData")
library(plotly)
load("/Users/pplsuser/Dropbox/phd/selection_replypaper/cletalk/plotlys.RData")
load(file="~/Desktop/korpused/wright-fisher_simulations4.RData")
source("/Users/pplsuser/Dropbox/phd/selection_replypaper/fitmap_plotters.R")

```

# A bit of background

---

# A bit of background
--

- All natural languages change over time
--

- Many have suggested that language change, like other evolutionary processes, involves both directed selection as well as stochastic drift .small[(Sapir1921, Jespersen1922, Andersen1987, Mcmahon1994, Croft2000, Blythe2012)] 
- Number of ways in which selective biases may influence language change .small[(Kirby2008, Smith2013, Enfield2014, Croft2000, Haspelmath1999, Labov2011, Mcmahon1994, Zipf1949, Baxter2006, Daoust2017; +et-al.'s )]
--

- Signatures of selection should be inferable from the usage data .small[(Sindi2016, Reali2010, Bentley2008, Amato2018, Kander2017; +et-al.'s)]
---

background-image: url(papershot.png)
background-size: contain

---
# Newberry et al. 2017, Detecting evolutionary forces in language change

- _"...we quantify the strength of selection relative to stochastic drift in language evolution."_

--

- _"...time series derived from large corpora of annotated texts"_ 
   - English verb (ir)regularization; COHA
   - Frequency Increment Test (FIT)

--

- _"...this work provides a method for testing selective theories of language change against a null model and reveals an underappreciated role for stochasticity in language evolution."_

---

# The Frequency Increment Test (FIT)

- Feder et al. 2014 .small[(from a family of tests of selection, cf. refs in paper)]
- Series of relative variant frequencies $v_i \in (0,1)$ at time $t_i$ 
- Transformed into frequency increments 
- $Y_i = (v_i-v_{i-1}) / \sqrt{ 2v_{i-1}(1-v_{i-1})(t_i-t_{i-1}) }$
--

- Rationale: under neutral evolution, the increments $v_i-v_{i-1}$ are normally distributed with a mean of 0, and variance ~ $v_{i-1}(1-v_{i-1})(t_i-t_{i-1})$ (inversely proportional to effective population size; when $0<<v_i<<1$; Gaussian approximation of the Wright-Fisher diffusion process)
--

- Test under the null hypothesis of drift ~ test that the increments are normally distributed with a mean of 0 (e.g.: one-sample $t$-test).

---

```{r fitexample, eval=T, fig.height=3, fig.width=5, dpi=200}
#knitr::knit_exit() # titleslide debug handbrake

ex = 1/(1+exp(-21*(seq(0.1,0.9,length.out = 9)-0.5))); ex[8]=ex[8]-0.1
library(vioplot) # uses dev v0.3
  par(mar=c(2, 3, 1, 1), cex.axis=0.8)
 j=1
    #if(j==3) par(mar=c(3,2,0.5,0.2))
    plot(ex, type="n", 
         xlim=c(0.3,length(ex)+length(ex)/50),  
         ylim=c(-1,1.5), 
         xlab="", ylab="",xaxt = "n",yaxt="n",tck = 0.03)
    abline(h=c(0,1), lty=1, lwd=1, col="gray")
    if(length(ex)<25){ axis(1, 1:length(ex))} else { axis(1)}
    #axis(2,tck = 0.03, labels=NA)
    axis(4, labels=NA, tck=-0.03) # tck = 0.03,
    axis(2)

    v=ex; v=ifelse(v <= 0, 0.001, ifelse(v >= 1, 0.999, v))
    t=1:length(v)
    Y = rep(0,(length(v))); Y[1]=NA
    for (i in c(2:length(v))) { # put increment value on the correct slot in Y to plot
      Y[i] = (v[i] - v[i - 1])/sqrt( 2*v[i-1]*(1 - v[i-1])*(t[i] - t[i-1]) )
    }
    swp=NA;fp=NA
    try({swp=shapiro.test(Y)$p.value})
    #swp=lillie.test(Y)$p.value
    try({ fp=FIT(ex) })
    pval=function(p){x= ifelse(p<0.001,"<0.001",paste0("=",p,collapse=""));return( bquote(italic(p)[FIT]~.(x)) )}
    sval=function(p){x= ifelse(p<0.001,"<0.001",paste0("=",p,collapse=""));return( bquote(italic(p)[SW]~.(x)) )}
    text(length(ex),c(1.48), pval(round(fp,3)), cex=1.4, adj=c(1,0.5))
    text(length(ex),c(1.2), sval(round(swp,3)), cex=1.4, adj=c(1,0.5))

    #points(rep(0.85, length(Y)), Y, cex=2, pch="-", col="darkgray") # increment rug
    vioplot(Y[!is.na(Y)], rectCol = F, col = "gray94", border=NA,add = T,
            at = (par('usr')[1]+1)/2, drawRect = F,side="both", wex=length(v)/7, lwd=0.5)
    segments(x0=(par('usr')[1]+1)/2, x1=(par('usr')[1]+1)/2,y0=min(Y,na.rm=T),y1=max(Y,na.rm=T), col="gray",lend=2,lwd=0.4)
    vioplot(Y[!is.na(Y)], rectCol = F, col = NA, border="gray58",add = T,
            at = (par('usr')[1]+1)/2, drawRect = F,side="both", wex=length(v)/7, lwd=0.5)

    lines(ex, type="o", pch=20, lwd=1)  # actual values
    lines(Y, type="b", pch="-",cex=1.5, font=2, lwd=1, lty=3, col=rgb(0.3,0.3,0.3,0.8),font=2) # increment values

```

---

# Problem: how to bin the data for time series

- Microbial experiments: samples that are taken at chosen intervals and resequenced
- Common approach in corpora usage: bin fixed length time segments
    - there is always a minimal time precision threshold (COHA: years)
    - but often not enough observations at fine precision 
    - so: decades, years, days, minutes
    - example: daily newspaper

--

- Newberry et al.: use variable width quantile binning, n(bins) = log(total frequency). Assures ~same number of occurrences per bin (but bins cover different lengths of time)

---
.pull-left[ 
```{r, fig.width=5.5, fig.height=5,dpi=100}
p1
```
]
.pull-right[
```{r, fig.width=5.5, fig.height=5,dpi=100}
p2   
```
]

<span style="font-weight: bold;color: #D58639"> $p<0.05$ </span><br>
<span style="font-weight: bold;color: #CDCD8D"> $p<0.2$ </span><br>
<span style="font-weight: bold;color: #C1EFFF"> $p>0.2$ </span>

---
class:inverse
# Replication of Newberry et al. 2017 (36 verbs)
---
# Replication of Newberry et al. 2017 (36 verbs)

```{r, eval=T, fig.height=4, fig.width=9, dpi=500}
verbmap()
```
---

# Replication of Newberry et al. 2017 (36 verbs)

```{r, eval=T, fig.height=4, fig.width=9, dpi=500}
verbmap(match( c("light", "smell", "sneak", "wake"), opts[[1]][,1]))
```

---

# Replication of Newberry et al. 2017 (36 verbs)

```{r, eval=T, fig.height=4, fig.width=9, dpi=500}
verbmap(c(6:7,9:14,26))
```

---

# Replication of Newberry et al. 2017 (36 verbs)

```{r, eval=T, fig.height=4, fig.width=9, dpi=500}
verbmap()
```
---

# Some thoughts

--

- In broad strokes, the generalization by Newberry et al. 2017 holds - selection is indeed detected in only ~3..7 verbs (depending on binning), and drift is quite prevalent (at $\alpha=0.05$).

--

- However, for most individual time series, the FIT result varies between binnings (except for ~3 almost unambiguous cases)

--

- So is it a good approach to study language change?<br>Depends on the goal.

--

- But still, what's the deal with the variation in the results...?

---
```{r,eval=T, fig.height=4, fig.width=9, dpi=500}
doexampleplot(1:10)
```
.small[What's going on?]
---
```{r,eval=T, fig.height=4, fig.width=9, dpi=500}
doexampleplot(1:2)
```
.small[(e.g. _spill, burn_)]
---
```{r,eval=T, fig.height=4, fig.width=9, dpi=500}
doexampleplot(3:4)
```
.small[(e.g. _knit_)]
---
```{r,eval=T, fig.height=4, fig.width=9, dpi=500}
doexampleplot(5:6)
```
.small[(differences between number of bins)]
---
```{r,eval=T, fig.height=4, fig.width=9, dpi=500}
doexampleplot(7:8)
```
---
```{r,eval=T, fig.height=4, fig.width=9, dpi=500}
doexampleplot(9:10)
```
.small[(e.g., _tell_)]
---
class:inverse
# Simulating change and applying binning<br>to determine the reasonable application range<br>of the FIT
---

# Simulating change and binning
--

- Run a large number of Wright-Fisher simulations with 200 different selection coefficients $s \in [0,5]$

--

- 200 generations, the "mutant" starting at 5% and 50% of the population of size 1000.

--

- For each $s$, bin the series in successively fewer number of bins <br>e.g. 200 (bin length 1) -> 100 (length 2) -> 66 (length 3) etc

--

- Repeat every combination 100x for good measure

---

.pull-left[ 
```{r, fig.width=5.5, fig.height=5,dpi=100}
wf0
```
]
.pull-right[
```{r, fig.width=5.5, fig.height=5,dpi=100}
wf  
```
]

---

```{r,eval=T, fig.height=4.5, fig.width=9, dpi=500}
doslideplot()
```

---
```{r,eval=T, fig.height=4.5, fig.width=9, dpi=500}
dofitmap(2)
```
.small[(start at 5%)]
---
```{r,eval=T, fig.height=4.5, fig.width=9, dpi=500}
dofitmap(1)
```
.small[(start at 50%)]
---

# Observations

- The FIT is insensitive to binning when selection is too weak ( $s<0.01$) to be detected; beyond about $s>0.02$ (depending on the start value) sensitivity to binning increases (false negatives)
--

- $0.01<s<0.02$ is relatively insensitive; but also where binning can instead decrease the FIT $p$-value (false positives)
--

- The normality assumption is systematically violated when $s$ approaches 0.1 (unless extreme binning is applied, which increases the false negative rate)

---

# Range of applicability of the FIT for linguistic data

- Conditions where the FIT is not reliably applicable:
    - partially completed changes, too short series
    - too few data points (sensitive to binning & absorption adjustment)
    - too long series (multiple events or processes)
    - too high selection (particularly with high binning)
    - small near-boundary fluctuations (false positives)
    - steep changes from boundary->non-boundary values
    - monotonically increasing series (normality assumption)
- Where it is:
    - weak selection, non-monotonic series away from 0/1, but window covering enough of (a single) change

---
class: inverse

# Conclusions
--

- What a time to be alive! .small[(data, methods, tools)]
--

- We evaluated the proposal of Newberry et al. 2017<br>Found that the results are dependent corpus binning, small sample effects, and the specifics of the FIT. 
- Testing vs generating hypotheses; degrees of freedom
--

- Fixing the issues would invite answers to numerous interesting questions

---
class: inverse

- Fixing these issues would invite answers to numerous interesting questions such as
--

   - Do different parts of grammar/lexicon experience stronger drift? 
--

   - What is the relationship of selection strength and niche in language change? .small[(cf. Laland2001, Altmann2011)] 
--

   - Can different types of selection (top-down, grassroots, momentum) be distinguished?  .small[(Amato2018, Stadler2016)]
--

   - What is the role of drift in creole evolution? .small[(Strimling2015)]
--

   - In semantic change? .small[(Hamilton2016)]
--

   - Are some languages changing more due to drift than others? Relation to community size? .small[(Reali2018, Atkinson2015)<br>(+et-al.'s)]

---
class: inverse
# Conclusions

- What a time to be alive! .small[(data, methods, tools)]
- We evaluated the proposal of Newberry et al. 2017<br>Found that the results are dependent corpus binning, small sample effects, and the specifics of the FIT. 
- Testing vs generating hypotheses; degrees of freedom
- Fixing the issues would invite answers to numerous interesting questions
- Identifying the role of drift vs selection in language change is an important goal, but: care with applying such tests to linguistic data, to avoid biases due to specifics of the domain and the particular test.
- Slides, code & arXiv link at http://andreskarjus.github.io

---
class: inverse
# Acknowledgements...

- Kenny Smith, Richard Blythe, Simon Kirby
- Mitchell Newberry
- Alison Feder
- .small[Support by the Kristjan Jaak program, funded by the Archimedes Foundation & Ministry of Education and Research of Estonia]