### 2025-01-05 news.Rmd

---
title: ''
output: html_document
theme: spacelab
---
<head>
<base target="_blank">
</head>
<style>
.main-container {
  padding: 0px;
}
ul {
  margin: 0; /* Removes any default margin */
  padding-left: 15px; /* Moves bullets slightly to the left */
  list-style-position: outside; /* Keeps bullets outside the text block */
  line-height: 1.1; /* Reduces the space between lines */
}

</style>

<div style="width: 100%; padding:0px; indent-left:0px">

- 2024: managed to publish 5 papers (3 preprints still in press) and present in a handful of very nice conferences.
- Mentored at the EdTech 2024 in Tallinn again. Also numerous AI workshops in Estonian schools this year.
- Got awarded the Nationally recognized science communicator quality mark at this year's [awards](https://etag.ee/eesti-teaduse-populariseerimise-elutoopreemia-palvis-toivo-maimets/)
- Visited and gave talks at the Meertens Institute and ILLC in Amsterdam
- Visited the DISSINET group in Brno to discuss applications of AI to historical data
- Appointed lecturer in DH and AI at Tallinn University (September 2024)
- Gave a workshop at the Baltic DH summer school 2024 in Riga again
- Participated at the UNESCO Global AI Ethics Forum (more on this in this [LinkedIn post](https://www.linkedin.com/posts/andreskarjus_aiethics-ai-activity-7159597653553524736-VnZv?utm_source=share&utm_medium=member_desktop) and Twitter/X [threads](https://twitter.com/AndresKarjus)) (February 2024)
- Meanwhile, have done more AI workshops for schools and university departments across Estonia (details&booking requests see [here](datafigure.eu))
- Now out in Corpus Linguistics and Linguistic Theory: "Reliable identification of selection mechanisms in language change" | [Open access](https://doi.org/10.1515/cllt-2023-0064)
- Mentored at the EdTech [Hackathon](https://eventornado.com/event/edtech-hack-2023?s=1#home) 2023 in Tallinn
- Joined Estonian Business School as a senior research fellow, initially part-time; continuing also at Tallinn University.
- New preprint [Machine-assisted mixed methods: augmenting humanities and social sciences with artificial intelligence](https://arxiv.org/abs/2309.14379) (24.09.2023)
- Two new preprints this week: ["Evolving Linguistic Divergence on Polarizing Social Media"](https://arxiv.org/abs/2309.01659) and ["A Framework for the Analysis of Historical Newsreels"](https://doi.org/10.31235/osf.io/a4xsp)  (06.09.2023)
- Out now in Frontiers Neuroscience: Yilmaz, Lotman, Karjus, Tikka, "An embodiment of the cinematographer: emotional and perceptual responses to different camera movement techniques" | [open access](https://doi.org/10.3389/fnins.2023.1160843) (03.07.2023)
- Out now in EPJDS: Karjus et al., "Compression ensembles quantify aesthetic complexity and the evolution of visual art" | [open access](https://doi.org/10.1140/epjds/s13688-023-00397-3) | [code](https://github.com/andreskarjus/compression_ensembles) (14.06.2023)
- Published an opinion piece in the popular science outlet [Novaator](https://novaator.err.ee/1609007981/andres-karjus-tehisintellekt-seab-juba-sel-sugisel-koolid-sundvaliku-ette) on how generative AI will affect education in the near future (in Estonian; 16.06.2023)
- New preprints: Ohm et al. "Collection Space Navigator: An Interactive Visualization Interface for Multidimensional Datasets", and Zemaityte et al. "Quantifying the global film festival circuit: Networks, diversity, and public value creation"
- Spring 2023: Been doing more workshops at multiple Estonian higher education institutions, schools and companies on generative AI, text and image generators; several workshops and public speaking engagements coming up; more info at [datafigure.eu]((https://datafigure.eu)
- Gave a talk about experiences of teaching digital skills, programming in R and data visualization, at the "Cross-university collaboration in Digital Humanities & Social Science (DHSS) and Digital Humanities & Cultural Heritage (DHCH) Education" workshop of the DHNB2023 conference | [slides](https://datafigure.eu/reports/AndresKarjusDatafigure_CrossUniversityDHworkshop.pdf)
- Been doing a series of [workshops](https://twitter.com/AndresKarjus/status/1625844088023969792) at Tallinn University on the topic of ChatGPT, generative AI, what's coming and how it will be/is already affecting teaching and learning in the university and beyond. (15.02.2023)
- New preprint "Quantifying public value creation by public service media using big programming data" [socarxiv](https://osf.io/preprints/socarxiv/hysma/)

- Talk "Exploring Estonian Public Television Production 2004-2020 Using Big Programming Data" at the 8th Estonian Digital Humanities [Conference](https://dh.org.ee/events/dhe2022) (05.10.2022)

- We received an applied research prize from the Tallinn city government for our industry collaboration project with Ekspress Group (28.09.2022) | [News piece in Estonian](https://www.tlu.ee/bfm/uudised/rakenduslike-teadustoode-konkursi-voitis-indrek-ibruse-mark-metsa-ja-andres-karjuse)

- Poster "Evolving Linguistic Divergence in Socio-Political Polarities", at the [JCoLe](https://sites.google.com/view/joint-conf-language-evolution) Joint Conference on Language Evolution in Kanazawa, Japan (August 2022)

- This summer I'm teaching R and data visualization workshops in 2 [summer schools](https://andreskarjus.github.io/artofthefigure/), in Latvia and Czech Republic.

- Poster "Linguistic divergence in American English along socio-political polarities", at the [IC2S2](https://www.ic2s2.org/) Computational Social Science Conference in Chicago (20.07.2022) | [pdf](https://andreskarjus.github.io/talks/ic2s2/IC2S2_poster_Linguistic_divergence_in_American_English_along_socio-political_polarities.pdf)

- New preprint: "Compression ensembles quantify aesthetic complexity and the evolution of visual art" [arxiv](https://arxiv.org/abs/2205.10271)

- Invited talk on language change and big social media data, in the research seminar series of the School of Humanities of Tallinn University (23.02.2022)

- Invited talk on linguistic corpora and experiments (and media&culture data) at the seminar of the Poncelet laboratory in Moscow (17.11.2021)

- Talk at the [Conference on Complex Systems 2021](https://ccs2021.univ-lyon1.fr/) in Lyon: "Capturing aesthetic complexity in art using compression ensembles" (29.10.2021) | [slides here](https://andreskarjus.github.io/talks/ccs2021/Karjus_CCS2021_Capturing_aesthetic_complexity.pdf)

- Now out in [Cognitive Science](https://doi.org/10.1111/cogs.13035) and also presented at Protolang 7 - "Conceptual similarity and communicative need shape colexification: an experimental study" (08.09.2021) | [pdf](https://andreskarjus.github.io/papers/Karjus_2021__Conceptual_Similarity_and_Communicative_Need_Shape_Colexification_An_Experimental_Study__Cognitive_Science.pdf) | [Protolang slides](https://andreskarjus.github.io/talks/andres_karjus_protolang7.pdf) | [tweet-summary](https://twitter.com/AndresKarjus/status/1435536919459418113)

- Taught a workshop on using R for data exploration and visualization, and moderated a research methods troubleshooter workshop, at the [Baltic Summer School of Digital Humanities 2021](https://www.nlib.ee/et/digital-humanities-tallinn-2021) (august 2021)

- Presented a [poster "Capturing aesthetic complexity in art
using compression ensembles"](https://andreskarjus.github.io/talks/cultureconference2021/) (click for pdf) at [Culture Conference 2021](https://culture-conference.com/)

- Gave a talk on my recent PhD and ongoing postdoc research in the [TÜling](https://www.keel.ut.ee/et/tuling) lecture series (27.04.2021)

- New preprint: "Conceptual similarity and communicative need shape colexification" | [preprint](arxiv.org/abs/2103.11024) |  [tweet-summary](https://twitter.com/AndresKarjus/status/1374332704662167557)

- Gave an invited talk about my PhD research on lexical dynamics at the [Colloquium for Computational Linguistics and Linguistics in Stuttgart](https://www.ims.uni-stuttgart.de/en/research/colloquium/). (11.01.2021) [Slides](https://andreskarjus.github.io/talks/stuttgart2021/andres_karjus_lexical_dynamics_slides_stuttgart2021.pdf) | [Recording](https://www.ims.uni-stuttgart.de/documents/aktuelles/veranstaltungen/210111-Karjus-Talk-Colloq-Recording.mp4)

- With the corrections approved and paperwork competed, seems I now have a PhD (26.11.2020)

- Was happy to host a job shadow (töövari) student at the CUDAN lab, via the Tagasi Kooli programme (16.11.2020)

- Defended my PhD [thesis](https://andreskarjus.github.io/thesis/Andres_Karjus_phd_thesis.pdf), "Competition, selection and communicative need in language change" and passed with minor corrections (29.10.2020)

- Started a postdoc at [CUDAN](http://cudan.tlu.ee/) lab at Tallinn University, provisionally as a visiting fellow until my defense (01.10.2020)

- Happy to have been invited to teach [data visualization](https://andreskarjus.github.io/artofthefigure/) and R at the [Digital Methods in Humanities and Social Sciences](https://digitalmethods.ut.ee/) Summer School (26.08.2020)

- Delivered another dataviz workshop for the University of Edinburgh Centre for Data, Culture and Society (27.07.2020)

- New [preprint](https://arxiv.org/abs/2006.09277) out, Communicative need modulates competition in language change

- Our "Challenges in detecting evolutionary forces in language change using diachronic corpora" now [published in Glossa](http://doi.org/10.5334/gjgl.909) (07.05.2020)

- Gave an online CLE talk about my recent PhD research on lexical competition and colexification (31.03.2020) | [slides](https://andreskarjus.github.io/cletalk2020) | [recording](https://youtu.be/-utY5ERe6OY)

- I was co-organizing a workshop titled *Drift, selection, change, and shift* that was supposed to take place at Evolang2020 in April. Unfortunately the conference got (understandably) cancelled due to the world being on fire. We had a great lineup of speakers planned though [(workshop homepage here)](https://driftselection-evolang2020.github.io/) and hope to reconvene next year.

- My first PhD paper (that's been out as a preprint since summer 2018) is now published in Language Dynamics and Change: [Karjus, A., Blythe, R. A., Kirby, S., & Smith, K. (2020). Quantifying the dynamics of topical fluctuations in language](https://doi.org/10.1163/22105832-01001200)

---

- Delivered another dataviz workshop, this time for the University of Edinburgh Centre for Data, Culture and Society (01.11.2019)

- Presenting a talk at the [RUSE 2019](http://rusesymposium.org.uk/), the Symposium on Representations, Usage and Social Embedding in Language Change (a slightly updated version of the CL2019 talk). Slides [here](https://andreskarjus.github.io/ruse2019).

- Presented a talk on lexical competition at the [CL2019](http://www.cl2019.org/). Slides with interactive plots [here](https://andreskarjus.github.io/cl2019talk.html). Also teaching a dataviz [workshop](https://andreskarjus.github.io/artofthefigure/cl2019).

- Presenting an updated version of the communivative need & lexical competition poster at the [Culture Conference 2019](https://culture-conference.com/programme-2019/) in Stirling, 1-2 July 2019. Online version with interactive plots [here](https://andreskarjus.github.io/lexcom_poster). More languages this time!

- I'll be teaching workshops at the following venues this summer: [SCOSYA Data Hack](https://scotssyntaxatlas.ac.uk/workshops/data-hack/), [SGSSS Summer School](https://www.sgsss.ac.uk/summer-school-2019/), [SGSAH Summer School](https://www.sgsah.ac.uk/news/summerschool/), [Digital Humanities Conference 2019](https://dh2019.adho.org/), [International Corpus Linguistics Conference 2019](http://www.cl2019.org/). All materials available [here](https://andreskarjus.github.io/artofthefigure/).

- Taught an aRt of the Figure series workshop for the Corpus Linguistics in Scotland network (12.04), titled [Data visualization with R for corpus linguists](https://andreskarjus.github.io/artofthefigure)

- Edinburgh Language Lunch 11.04: (re-)presented the "Modelling lexical interactions in diachronic corpora" poster, originally presented at the [Inaugural ISLE workshop](https://www.comparativelinguistics.uzh.ch/en/events/ISLE-inaugural-workshop.html), here is an [online version with interactive plots](https://andreskarjus.github.io/isle_poster) and an (updated!) appendix.

- The Language Evolves school workshops are [finished for 2018/2019](https://twitter.com/kennysmithed/status/1113456451815661568), with Hermitage Park Primary being the last one. Happy to have been part of it.

- Poster to be presented at the [Inaugural ISLE workshop](https://www.comparativelinguistics.uzh.ch/en/events/ISLE-inaugural-workshop.html) in April, here is an online [version with interactive plots](https://andreskarjus.github.io/isle_poster).

- Entered [a video](https://www.youtube.com/watch?v=t8InfFEIINc) to Science Magazine's 11th [Dance Your Ph.D contest](https://www.sciencemag.org/projects/dance-your-phd). The idea is to explain your research topic to the general public though interpretive dance.

---

- [Slides](https://andreskarjus.github.io/evoforces_cletalk/slides.html) of the talk _Challenges in detecting evolutionary forces in language change using diachronic corpora_ presented at the [CLE](http://www.lel.ed.ac.uk/cle/) seminar series, 06.11.2018 | paper preprint now also on [arXiv](http://arxiv.org/abs/1811.01275) | [code](https://github.com/andreskarjus/wfsim_fit)

- Out now in Language Dynamics and Change: Karjus & Ehala 2018, Testing an agent based model of language choice on sociolinguistic survey data | [Online at Brill](https://doi.org/10.1163/22105832-00802004) | [PDF](Karjus_Ehala_2018_Testing_an_agent-based model_of_language_choice_LDCpostprint.pdf) | [dataset](http://doi.org/10.15155/1-00-0000-0000-0000-00162L): representative sample of 1000 respondents from Estonia, 200 questions on language attitudes and use

- "Two problems and solutions in evolutionary corpus-based language dynamics research" to be presented @ SLE 2018: [slides](https://andreskarjus.github.io/sle_tallinn_languagedynamics) with interactive plots and [code](https://github.com/andreskarjus/sle_tallinn_languagedynamics) now available.

- July 2018: the [Quantifying the dynamics of topical fluctuations in language](https://arxiv.org/abs/1806.00699) paper got awarded the University of Edinburgh PPLS [Gray postgraduate publication prize](https://www.ed.ac.uk/ppls/philosophy/current/postgraduate/fees-and-funding/ppls-gray-postgraduate-publication-prize).

- Code to run the topical-cultural advection model on COHA (or other corpora; as described in the recent [arXiv preprint](https://arxiv.org/abs/1806.00699)) is now available [on Github](https://github.com/andreskarjus/topical_cultural_advection_model).

- _Selection above the baseline: quantifying the advection effect in four domains of cumulative culture_ @ the [Applications in Cultural Evolution](https://cultevol.ut.ee/) conference: slides with a bunch of interactive plots that allow playing with the massive datasets [now available here](https://andreskarjus.github.io/cultevol_tartu_slides), and all the code [here](https://github.com/andreskarjus/cultural_advection_Tartuce).

</div>